### 2013-12-31 pkg/inst/doc/msm-manual.Rnw
Make trivial vignette to test build
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/msm@35 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

%%\VignetteIndexEntry{User guide to msm with worked examples}
%\VignettePackage{msm}
\documentclass{article}

\begin{document}

\maketitle

\begin{abstract}

Test

\end{abstract}

\tableofcontents
\clearpage

\section{Multi-state models}
\label{sec:multistate}
Test

\end{document}