### 2013-12-31 pkg/inst/doc/msm-manual.Rnw
Test build again: just preamble, title and abstract
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/msm@37 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

%%\VignetteIndexEntry{User guide to msm with worked examples}
%\VignettePackage{msm}
\documentclass{article}

%% Need to modify Sweave.sty to pass pdftex option to graphicx.
\usepackage{Sweave-local}
\usepackage{times}
\usepackage{url}
\addtolength{\textwidth}{2cm}

\newcommand{\Exp}{\mathop{\mathrm{Exp}}}
\newcommand{\etal}{{\textit{et al.}}}
\newcommand{\Rfunction}[1]{{\texttt{#1}}}
\newcommand{\Robject}[1]{{\texttt{#1}}}
\newcommand{\Rpackage}[1]{{\textit{#1}}}
\newcommand{\Rmethod}[1]{{\texttt{#1}}}
\newcommand{\Rfunarg}[1]{{\texttt{#1}}}
\newcommand{\Rclass}[1]{{\textit{#1}}}

%% version <- 1.3

<<echo=FALSE>>=
version <- gsub("Version: +", "",
               packageDescription("msm", lib.loc="../../..")$Version)
@

\title{Multi-state modelling with R: the {\tt msm} package \vskip 0.2in
  \large{Version %
<<echo=FALSE,results=tex>>=
cat(version)
@
\vskip 0.1in
<<echo=FALSE,results=tex>>=
cat(format(Sys.time(), "%d %B, %Y"))
@
}}
\author{Christopher Jackson\\MRC Biostatistics Unit\\Cambridge, U.K.\\ \texttt{chris.jackson@mrc-bsu.cam.ac.uk}}
\date{}
%% label with date when Rnw is compiled, not when tex is compiled. should be release date of package

\bibliographystyle{unsrt}

\begin{document}

\maketitle

\begin{abstract}

  The multi-state Markov model is a useful way of describing a process
  in which an individual moves through a series of states in
  continuous time.  The \Rpackage{msm} package for R allows a general
  multi-state model to be fitted to longitudinal data. Data often
  consist of observations of the process at arbitrary times, so that
  the exact times when the state changes are unobserved.  For example,
  the progression of chronic diseases is often described by stages of
  severity, and the state of the patient may only be known at doctor
  or hospital visits.  Features of \Rpackage{msm} include the ability
  to model transition rates and hidden Markov output models in terms
  of covariates, and the ability to model data with a variety of
  observation schemes, including censored states.

  Hidden Markov models, in which the true path through states is only
  observed through some error-prone marker, can also be fitted.  The
  observation is generated, conditionally on the underlying states,
  via some distribution.  An example is a screening misclassification
  model in which states are observed with error.  More generally,
  hidden Markov models can have a continuous response, with some
  arbitrary distribution, conditionally on the underlying state.

  This manual introduces the theory behind multi-state Markov and
  hidden Markov models, and gives a tutorial in the typical use of the
  \Rpackage{msm} package, illustrated by some typical applications to
  modelling chronic diseases.

  Much of the material in this manual is published, in a more concise
  form, in Journal of Statistical Software (2011) 38(8):1-29,
  \url{http://www.jstatsoft.org/v38/i08/}

\end{abstract}

\section{Multi-state models}
\label{sec:multistate}
Test

\end{document}